|

### 2022-06-22 docs/rmd/paper.Rmd
rendered doc

---
title: |
  <center>Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression: </center>
  <center>
    A comparison of $\epsilon$-Lexicase Selection and Tournament Selection
  </center>
  <center>--------------------------------------------------------</center>
  <center>Student: Roman Höhn</center>
  <center>Date of Birth: 1991-04-14</center>
  <center>Place of Birth: Wiesbaden, Hesse</center>
  <center>Student ID: 2712497</center>
  <center>Supervisor: David Wittenberg</center>
  <center>--------------------------------------------------------</center>
  <center>Research Project for Seminar Information Systems (03.996.3299)</center>
  <center>Chair of Business Administration and Computer Science</center>
  <center>Johannes Gutenberg University Mainz</center>
  <center>Summerterm 2022</center>
  |
date: "Date of Submission: `r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["float"]
    fig_caption: yes
    number_sections: true
    includes:
      in_header: settings.tex
fontsize: 12pt
# toc: yes
# toc_depth: 4
bibliography: bib/bib/ref.bib
csl: bib/csl/harvard1.csl
# include-before: '`\newpage{} `{=latex}'
include-before: '`\thispagestyle{empty} `{=latex}'
---

@@

The first description^[@lexicase_first_desciption, p.4] of the lexicase parent selection algorithm by @lexicase_first_desciption is quoted verbatim below for reference:

@@

> $H0_{3}$: Size differences exist between the distribution underlying the samples produced by $\epsilon$-Lexicase and the distribution underlying samples of tournament selection

@@

In a final step I analyzed the growth behavior for both GP systems. Figure 3 visualizes the average size of individuals inside the population as well as the average size of the best performing individual for each generation of the evolution.

![Mean Size](./plots/size_combined.png)
... interpretation...

Again, a MWU test is conducted to test for statistical differences in the underlying distribution of the sizes measured during the experiment. The results are summarized in table 10.

@@

@@

To gain further insight into the differences in generalization behavior and to adress the research question with a higher level of certainty, the experiment I conducted should be repeated for many different GP configurations. Some exemplary parameters that could drastically influence the results of this experiment include the population size, the total number of generations, the inclusion of elitism or the computation of $\epsilon$ in lexicase selection. Other factors whose influence on generalization behavior could be important include different methods to compute an individuals fitness (e.g. mean absolute error or root mean squared error) and testing out different variations of the genetic operators for mutation and crossover.

@@

>>>>

---
title: |
  <center>Analyzing the influence of Selection on Genetic Programming's Generalization ability in Symbolic Regression: </center>
  <center>
    A comparison of $\epsilon$-Lexicase Selection and Tournament Selection
  </center>
  <center>--------------------------------------------------------</center>
  <center>Student: Roman Höhn</center>
  <center>Date of Birth: 1991-04-14</center>
  <center>Place of Birth: Wiesbaden, Hesse</center>
  <center>Student ID: 2712497</center>
  <center>Supervisor: David Wittenberg</center>
  <center>--------------------------------------------------------</center>
  <center>Research Project for Seminar Information Systems (03.996.3299)</center>
  <center>Chair of Business Administration and Computer Science</center>
  <center>Johannes Gutenberg University Mainz</center>
  <center>Summerterm 2022</center>
  |
date: "Date of Submission: `r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
    extra_dependencies: ["float"]
    fig_caption: yes
    number_sections: true
    includes:
      in_header: settings.tex
# indent: true
fontsize: 12pt
# toc: yes
# toc_depth: 4
bibliography: bib/bib/ref.bib
csl: bib/csl/harvard1.csl
# include-before: '`\newpage{} `{=latex}'
include-before: '`\thispagestyle{empty} \newpage{}`{=latex}'
---

@@

The first description of the lexicase parent selection algorithm is quoted verbatim below for reference [@lexicase_first_desciption, p.4]:

@@

> $H0_{3}$: No differences in program size exist between the distribution underlying the samples produced by $\epsilon$-Lexicase and the distribution underlying samples of tournament selection

@@

In a final step I analyze the growth behavior for both GP systems to determine if differences in program size can be observed based on selection operators. Figure 3 visualizes the average size of individuals inside the population as well as the average size of the best performing individual for each generation and both GP systems. 
![Mean Size](./plots/size_subplotted.png)

Both subplots display a very similar behavior that is common for GP, we observe a strong trend for solutions to grow in size with each additional generation that is concluded. Another observation  that can be made is that the best solutions of each generation almost always have a larger size than the population average. On first impression, the differences between both selection operators appears to be insignificant. Again, I conducted a MWU test to test for statistical differences in the underlying distribution of sizes measured during the experiment.

The resulting p-values of the MWU test are summarized in table 10.

@@

None of the differences in mean size are statistically significant. Again, based on the results obtained in my experiment, the evidence produced fails to reject the null hypothesis $H0_{3}$ that seletion has no significant effect on differences in program size.

@@

To gain further insight into the differences in generalization behavior and to adress the research question with a higher level of certainty, the experiment I conducted should be repeated for different GP configurations. Some exemplary parameters that could drastically influence the results of this experiment include the population size, the total number of generations, the inclusion of elitism or the computation of $\epsilon$ in lexicase selection. Other factors whose influence on generalization behavior could be important include different methods to compute an individuals fitness (e.g. mean absolute error or root mean squared error), testing different variations of the genetic operators for mutation and crossover and experimenting with different variations of the training:testing split size.

@@

```{zsh github, include=FALSE}
git add * && git commit --m "rendered doc" && git push
```
